### 2024-10-07 doc/Rogers_20XX/bat_guano.Rmd

---
title: A template for the *arxiv* style
authors:
  - name: Timothy J. Rogers
    thanks: 
    department: Department of Bioinformatics
    affiliation: University of North Carolina
    location: Charlotte, NC 28223
    email: troger50@charlotte.edu
  - name: Richard A. White
    department: Department of Bioinformatics
    affiliation: University of North Carolina
    location: Charlotte, NC 28223
    email: rwhit101@charlotte.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - guano
  - microbiome
  - bat
  - virome
  - metagenomics
bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE)`" 
#bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
---

# Introduction

Here goes an introduction text

# Headings: first level
\label{sec:headings}

You can use directly LaTeX command or Markdown text.

LaTeX command can be used to reference other section. See Section \ref{sec:headings}.
However, you can also use **bookdown** extensions mechanism for this.

## Headings: second level

You can use equation in blocks

$$
\xi _{ij}(t)=P(x_{t}=i,x_{t+1}=j|y,v,w;\theta)= {\frac {\alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}{\sum _{i=1}^{N} \sum _{j=1}^{N} \alpha _{i}(t)a^{w_t}_{ij}\beta _{j}(t+1)b^{v_{t+1}}_{j}(y_{t+1})}}
$$

But also inline i.e $z=x+y$

### Headings: third level

Another paragraph. [@deleon_etal18] and see bob.

# Examples of citations, figures, tables, references
\label{sec:others}

The documentation for \verb+natbib+ may be found at

You can use custom blocks with LaTeX support from **rmarkdown** to create environment.

::: {.center latex=true}
  <http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}>
:::

Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.

You can insert LaTeX environment directly too.

\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}

produces

\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}

## Figures

You can insert figure using LaTeX directly.

See Figure \ref{fig:fig1}. Here is how you add footnotes. [^Sample of the first footnote.]

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

But you can also do that using R.

```{r fig2, fig.cap = "Another sample figure"}
plot(mtcars$mpg)
```

You can use **bookdown** to allow references for Tables and Figures.

## Tables

Below we can see how to use tables.

See awesome Table~\ref{tab:table} which is written directly in LaTeX in source Rmd file.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

You can also use R code for that.

```{r}
knitr::kable(head(mtcars), caption = "Head of mtcars table")
```

## Lists

- Item 1
- Item 2 
- Item 3

### 2024-10-08 workflow/scripts/retreve_host_references.sh
Completed first section of snakemake scripts and ran test run

@@ -14,8 +14,22 @@
 # This code block dowloads the genomes for representative #
 # sequences using the datasets program.                   #
 ###########################################################
-# datasets download genome accession --inputfile snakemake_input[accessions] \
-#     --include gff3,rna,cds,protein,genome,seq-report
+#### Calling in input and output from the rule doc:
+input=(${snakemake_input[accessions]})
+output=(${snakemake_output[ncbi_zip]})
 
-datasets download genome accession --inputfile /projects/raw_lab/projects/Bats/bat_gut_micro/data/references/bat_host_accessions.txt \
-    --include gff3,rna,cds,protein,genome,seq-report
+#### Running datasets to retrieve sequences
+datasets download genome accession --inputfile "$input" \
+    --include gff3,rna,cds,protein,genome,seq-report \
+    --filename "$output"
+
+
+
+
+##########################################################
+# This code chunk was originally used to test the script #
+##########################################################
+# datasets download genome accession --inputfile /projects/raw_lab/projects/Bats/bat_gut_micro/data/references/test_bat_host_accessions.txt \
+#     --include gff3,rna,cds,protein,genome,seq-report --filename /projects/raw_lab/projects/Bats/bat_gut_micro/test/bat_references.zip
+
+# unzip /projects/raw_lab/projects/Bats/bat_gut_micro/test/bat_references.zip /projects/raw_lab/projects/Bats/bat_gut_micro/test/

### 2024-10-08 workflow/scripts/retreve_host_references.sh
Updates

@@ -15,13 +15,12 @@
 # sequences using the datasets program.                   #
 ###########################################################
 #### Calling in input and output from the rule doc:
-input=(${snakemake_input[accessions]})
-output=(${snakemake_output[ncbi_zip]})
+
 
 #### Running datasets to retrieve sequences
-datasets download genome accession --inputfile "$input" \
+datasets download genome accession --inputfile $1 \
     --include gff3,rna,cds,protein,genome,seq-report \
-    --filename "$output"
+    --filename $2
 
 
 

### 2024-10-08 workflow/scripts/trimming.sh

@@ -3,7 +3,7 @@
 ################################################################
 # author: TJ Rogers                                            #
 # input: Name of the raw files (without the path)              #
-# output: places cleaned reads into data/processed/clean_reads #
+# output: places trimed reads into data/temp                   #
 #                                                              #
 # Trimming the raw reads.                                      #
 ################################################################
@@ -11,41 +11,32 @@
 ###################################################
 # Send all stderr from this script to the log file #
 ###################################################
-exec 2> "${snakemake_log[0]}" 
+# exec 2> "${snakemake_log[0]}" 
 
 #####################################
 # Call read files from the rule doc #
 #####################################
 # Raw
-raw_reads=(${snakemake_input[reads]})
-rr1="${raw_reads[0]}"
-rr2="${raw_reads[1]}"
+# raw_reads=(${snakemake_input[reads]})
+# rr1="${raw_reads[0]}"
+# rr2="${raw_reads[1]}"
 
-# Temp
-temp_reads=(${snakemake_output[tem_reads]})
-tr1="${temp_reads[0]}"
-tr2="${temp_reads[1]}"
 
-# Clean
-clean_reads=(${snakemake_output[final_reads]})
-cr1="${clean_reads[0]}"
-cr2="${clean_reads[1]}"
+# # Temp
+# temp_reads=(${snakemake_output[tem_reads]})
+# tr1="${temp_reads[0]}"
+# tr2="${temp_reads[1]}"
+
+# # Clean
+# clean_reads=(${snakemake_output[final_reads]})
+# cr1="${clean_reads[0]}"
+# cr2="${clean_reads[1]}"
 #####################
 # Removing adapters #
 #####################
-bbduk.sh -Xmx1g in1="$rr1" \
-    in2="$rr2" \
-    out1="$tr1" \
-    out2="$tr2" \
-    ref=adapters.fa ktrim=r k=21 \
-    qin=auto mink=11 hdist=2 tbo tpe
+bbduk.sh -Xmx1g in1=$1 in2=$2 out1=$3 out2=$4 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/adapters.fa ktrim=r k=21 qin=auto mink=11 hdist=2 tbo tpe
 
 #################
 # Removing phiX #
 #################
[…]

### 2024-10-08 workflow/scripts/host_read_removal.sh
Rerunning pipeline and constructing host_read_removal.sh script

@@ -1,0 +1,35 @@
+#!/usr/bin/env bash
+
+################################################################
+# author: TJ Rogers                                            #
+# input: Name of the trimmed files (without the path)          #
+# output: places cleaned reads into data/processed/clean_reads #
+#                                                              #
+# Removing host sequences.                                     #
+################################################################
+
+#### Create bowtie2 database of the host genome
+bowtie2-build host_genome_sequence.fasta host_DB
+
+#### Map reads against host sequence database, keeping both mapped and unmapped reads
+bowtie2 -p 8 -x host_DB \
+-1 SAMPLE_R1.fastq.gz \
+-2 SAMPLE_R2.fastq.gz \
+-S SAMPLE_mapped_and_unmapped.sam
+
+#### Convert sam files to bam files to save space
+samtools view -bS SAMPLE_mapped_and_unmapped.sam > SAMPLE_mapped_and_unmapped.bam
+
+#### Collect reads not mapped to host sequences
+samtools view -b -f 12 -F 256 \
+    SAMPLE_mapped_and_unmapped.bam \
+    > SAMPLE_bothReadsUnmapped.bam 
+
+#### Sort bam files of unmapped reads
+samtools sort -n -m 5G -@ 2 SAMPLE_bothReadsUnmapped.bam -o SAMPLE_bothReadsUnmapped_sorted.bam
+
+#### Separate the paired-end reads into separate fastq files
+samtools fastq -@ 8 SAMPLE_bothReadsUnmapped_sorted.bam \
+  -1 SAMPLE_host_removed_R1.fastq.gz \
+  -2 SAMPLE_host_removed_R2.fastq.gz \
+  -0 /dev/null -s /dev/null -n

### 2024-10-08 workflow/scripts/trimming.sh

@@ -11,32 +11,29 @@
 ###################################################
 # Send all stderr from this script to the log file #
 ###################################################
-# exec 2> "${snakemake_log[0]}" 
 
 #####################################
 # Call read files from the rule doc #
 #####################################
-# Raw
-# raw_reads=(${snakemake_input[reads]})
-# rr1="${raw_reads[0]}"
-# rr2="${raw_reads[1]}"
 
-
-# # Temp
-# temp_reads=(${snakemake_output[tem_reads]})
-# tr1="${temp_reads[0]}"
-# tr2="${temp_reads[1]}"
-
-# # Clean
-# clean_reads=(${snakemake_output[final_reads]})
-# cr1="${clean_reads[0]}"
-# cr2="${clean_reads[1]}"
 #####################
 # Removing adapters #
 #####################
-bbduk.sh -Xmx1g in1=$1 in2=$2 out1=$3 out2=$4 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/adapters.fa ktrim=r k=21 qin=auto mink=11 hdist=2 tbo tpe
+# Check if final output files exist
+if [[ -f "$5" && -f "$6" ]]; then
+  echo "Final files $5 and $6 already exist. Skipping this sample..."
+  exit 0
+fi
 
-#################
-# Removing phiX #
-#################
-bbduk.sh  -Xmx1g in1=$3 in2=$4 out1=$5 out2=$6 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/phix174_ill.ref.fa.gz k=27 hdist=1 qtrim=rl trimq=17 cardinality=t mingc=0.05 maxgc=0.95
+# Check if intermediate files exist
+if [[ -f "$3" && -f "$4" ]]; then
+    echo "Intermediate files $3 and $4 found. Running step 2 to remove phiX sequences..."
+    # Removing phiX
+    bbduk.sh -Xmx1g in1=$1 in2=$2 out1=$3 out2=$4 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/adapters.fa ktrim=r k=21 qin=auto mink=11 hdist=2 tbo tpe
+else
+    echo "Intermediate files not found. Running both steps to remove adapter and phiX sequences..."
+    # Removing adapters
[…]

### 2024-10-08 workflow/scripts/host_read_removal.sh
Updates

@@ -8,28 +8,38 @@
 # Removing host sequences.                                     #
 ################################################################
 
+#Assigning arguments:
+sample=$(basename ${$1} -sg_1.fastq.gz) #Grabing the sample name
+r1=$1                                   #Forward read input
+r2=$2                                   #Reverse read input
+fr1=$3                                  #Forward read output
+fr2=$4                                  #Reverse read output
+host_genome=$5                          #Host genome to index
+
 #### Create bowtie2 database of the host genome
-bowtie2-build host_genome_sequence.fasta host_DB
+bowtie2-build "$host_genome" host_DB
 
 #### Map reads against host sequence database, keeping both mapped and unmapped reads
 bowtie2 -p 8 -x host_DB \
--1 SAMPLE_R1.fastq.gz \
--2 SAMPLE_R2.fastq.gz \
--S SAMPLE_mapped_and_unmapped.sam
+-1 "$r1" \
+-2 "$r2" \
+-S "$sample"_mapped_and_unmapped.sam
 
 #### Convert sam files to bam files to save space
-samtools view -bS SAMPLE_mapped_and_unmapped.sam > SAMPLE_mapped_and_unmapped.bam
+samtools view -bS "$sample"_mapped_and_unmapped.sam > "$sample"_mapped_and_unmapped.bam
 
 #### Collect reads not mapped to host sequences
 samtools view -b -f 12 -F 256 \
-    SAMPLE_mapped_and_unmapped.bam \
-    > SAMPLE_bothReadsUnmapped.bam 
+    "$sample"_mapped_and_unmapped.bam \
+    > "$sample"_bothReadsUnmapped.bam 
 
 #### Sort bam files of unmapped reads
-samtools sort -n -m 5G -@ 2 SAMPLE_bothReadsUnmapped.bam -o SAMPLE_bothReadsUnmapped_sorted.bam
+samtools sort -n -m 5G -@ 2 "$sample"_bothReadsUnmapped.bam -o "$sample"_bothReadsUnmapped_sorted.bam
 
 #### Separate the paired-end reads into separate fastq files
-samtools fastq -@ 8 SAMPLE_bothReadsUnmapped_sorted.bam \
-  -1 SAMPLE_host_removed_R1.fastq.gz \
-  -2 SAMPLE_host_removed_R2.fastq.gz \
[…]

### 2024-10-09 workflow/scripts/retreve_host_references.sh
Testing scripts

@@ -21,10 +21,7 @@
 datasets download genome accession --inputfile $1 \
     --include gff3,rna,cds,protein,genome,seq-report \
     --filename $2
-
-
-
-
+    
 ##########################################################
 # This code chunk was originally used to test the script #
 ##########################################################

### 2024-10-10 snakemake_project/workflow/scripts/retreve_host_references.sh
Updating working directory layout so different projects can be septerated

@@ -15,17 +15,25 @@
 # sequences using the datasets program.                   #
 ###########################################################
 #### Calling in input and output from the rule doc:
-
-
+accession=$1
+ncbi_zip=$2
+ncbi_unziped=$3
+refdir=$4
 #### Running datasets to retrieve sequences
-datasets download genome accession --inputfile $1 \
+datasets download genome accession --inputfile "$accession" \
     --include gff3,rna,cds,protein,genome,seq-report \
-    --filename $2
-    
+    --filename "$ncbi_zip"
+
+unzip "$ncbi_zip" -d "$refdir"
+
+mv "$refdir"ncbi_dataset/data/GCA_036768555.1/GCA_036768555.1_ASM3676855v1_genomic.fna "$refdir"ncbi_dataset/data/GCA_036768555.1/GCA_036768555.1.fna
+mv "$refdir"ncbi_dataset/data/GCA_028538775.1/GCA_028538775.1_mMyoYum1.0.hap1_genomic.fna "$refdir"ncbi_dataset/data/GCA_028538775.1/GCA_028538775.1.fna
+mv "$refdir"/ncbi_dataset/data/GCA_004126475.3/GCF_004126475.2_mPhyDis1.pri.v3_genomic.fna "$refdir"ncbi_dataset/data/GCA_004126475.3/GCF_004126475.3.fna
+mv "$refdir"/ncbi_dataset/data/GCA_019186645.2/GCF_019186645.2_TTU_PhHast_1.1_genomic.fna "$refdir"ncbi_dataset/data/GCA_019186645.2/GCF_019186645.2.fna
 ##########################################################
 # This code chunk was originally used to test the script #
 ##########################################################
-# datasets download genome accession --inputfile /projects/raw_lab/projects/Bats/bat_gut_micro/data/references/test_bat_host_accessions.txt \
+# datasets download genome accession --inputfile "$refdir"/test_bat_host_accessions.txt \
 #     --include gff3,rna,cds,protein,genome,seq-report --filename /projects/raw_lab/projects/Bats/bat_gut_micro/test/bat_references.zip
 
 # unzip /projects/raw_lab/projects/Bats/bat_gut_micro/test/bat_references.zip /projects/raw_lab/projects/Bats/bat_gut_micro/test/

### 2024-10-10 titan_project/bin/scripts/titan_binning2.py
Constructing python script for microbial binning

@@ -1,0 +1,74 @@
+import os
+import subprocess
+import argparse
+from glob import glob
+
+###########################################
+# Make sure everything is set up properly #
+###########################################
+# The default length (-l) setting is 1000bp.
+l=1000
+
+#########################################
+# Aligning reads to make coverage files #
+#########################################
+# Coverage file(s) for the binning programs
+# I need to create coverage files for each binning program. This will be coverage
+# files of each assembly accross all metagenomes. This will allow for better bin-
+# ing.
+def align_reads():
+    print()
+
+def make_cov_files():
+    print()
+
+
+#################################################################################
+# Setting up functions to run each binning, refining, and dereplication program #
+#################################################################################
+# Function to run concoct
+def run_concoct(assembly_path, output_dir):
+    concoct_output_dir = os.path.join(output_dir, "concoct") #define concoct directory inside of ouput_dir
+    os.makedirs(concoct_output_dir, exist_ok=True)           #create the concoct output directory
+
+    cmd = f"""
+    mamba run -n concoct concoct \
+    --composition_file {assembly_path} \
+    --coverage_file {assembly_path}.cov \
+    -b {concoct_output_dir}
+    """
+    subprocess.run(cmd, shell=True, check=True)
+    print(f"CONCOCT binning completed for {assembly_path}")
+
+# Function to run metabat2
+def run_metabat2(assembly_path, output_dir):
+    metabat2_output_dir = os.path.join(output_dir, "metabat2")
+    os.makedirs(metabat2_output_dir, exist_ok=True)
+
+    cmd = f"""
+    mamba run -n metabat2 metabat2 \
+    -i {assembly_path} \
+    -o {metabat2_output_dir}
+    """
+    subprocess.run(cmd, shell=True, check=True)
+    print(f"Metabat2 binning completed for {assembly_path}")
+
+# Function to run maxbin2
+def run_maxbin2(assembly_path, output_dir):
+    maxbin2_output_dir = os.path.join(output_dir, "maxbin2")
+    os.makedirs(maxbin2_output_dir,exist_ok=True)
+
+    cmd = f"""
+    mamaba run -n maxbin2 maxbin2 \
+    -contig {assembly_path} \
+    -out {maxbin2_output_dir}/maxbin2_output \
+    -abund {assembly_path}.abund
+    """
+    subprocess.run(cmd, shell=True, check=True)
+    print(f"MaxBin2 binning completed for {assembly_path}")
+
+def run_bin_refinner():
+    print()
+
+def run_drep():
+    print()

### 2024-10-11 titan_project/bin/scripts/titan_binning2.py
Adding mamba envs

@@ -16,7 +16,45 @@ l=1000
 # I need to create coverage files for each binning program. This will be coverage
 # files of each assembly accross all metagenomes. This will allow for better bin-
 # ing.
-def align_reads():
+def align_reads(assembly_path, reads, output_dir):
+    os.makedirs(ouput_dir, exist_ok=True)
+
+    outR1 = f"map_{os.path.basename(reads[0])}"
+    outR2 = f"map_{os.path.basename(reads[1])}"
+    if not outR1.endswith(".gz"):
+        outR1 += ".gz"
+    if not outR2.endswith(".gz"):
+        outR2 += ".gz"
+    outfile = [os.path.join(path, outR1), os.path.join(path, outR2)] #Need to figure out what this is doing
+
+    # Iterate over each sample and map the reads
+        for sample in samples:
+            sample_name = sample['sample_name']
+            forward_reads = sample['forward']
+            reverse_reads = sample['reverse']
+
+            # Define output paths
+            sam_file = os.path.join(assembly_output_dir, f"{sample_name}.sam")
+            sorted_bam_file = os.path.join(assembly_output_dir, f"{sample_name}.sorted.bam")
+
+            # Step 1: Run BWA-MEM to map reads and generate SAM file
+            bwa_mem_cmd = f"""
+            mamba run -n bwa bwa mem {assembly} {forward_reads} {reverse_reads} > {sam_file}
+            """
+            subprocess.run(bwa_mem_cmd, shell=True, check=True)
+
+            # Step 2: Convert SAM to BAM and sort it
+            sam_to_bam_cmd = f"""
+            mamba run -n samtools samtools view -bS {sam_file} | mamba run -n samtools samtools sort -o {sorted_bam_file}
+            """
+            subprocess.run(sam_to_bam_cmd, shell=True, check=True)
+
+            # Step 3: Index the sorted BAM file
+            index_bam_cmd = f"mamba run -n samtools samtools index {sorted_bam_file}"
+            subprocess.run(index_bam_cmd, shell=True, check=True)
+
+            # Step 4: Remove intermediate SAM file
+            os.remove(sam_file)
     print()
 
 def make_cov_files():

### 2025-03-03 phasegenome_out/doc/paper/manuscript.rmd
Updating repository

---
title: Resolving gut microbiome networks within Chiropterans
authors:
  - name: Timothy J. Rogers
    department: Department of Bioinformatics
    affiliation: University of North Carolina
    location: Charlotte, NC 28223
    email: troger50@charlotte.edu
  - name: Laurel R. Yohe
    department: Department of Bioinformatics and Genomics
    affiliation: North Carolina Research Center (NCRC)
    location:  Kannapolis, NC 28081
    email:  lyohe1@charlotte.edu
  - name: Richard A. White ${^\dagger}$
    department: Department of Bioinformatics
    affiliation: University of North Carolina
    location: Charlotte, NC 28223
    email: rwhit101@charlotte.edu
abstract: |
  Enter the text of your abstract here.
keywords:
  - Chiropterans
  - microbiome
  - virome
  - metagenomics
bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE)`" 
#bibliography: references.bib
biblio-style: unsrt
output: rticles::arxiv_article
geometry: margin=1.0in
font-size: 11pt
header-include:
  -\usepapackage(helvet)
  -\renewcommand*\familydefault{\sfdefault}

---
${\dagger}$ To whom correspondence should be addressed

\newpage
```{r}
library(data.table)
library(tidyverse)
```
## Introduction
* Bats are known carriers of human associated pathogens
* The reason bats are bioreactors is not understood
* The diet of bats may contribute to the gut microbiota makeup
* Phage associated with these microbiota can benifit and hinder microbial populations and have an impack on bat immune responses
* Some ideas suggest viral tolerance is linked to 
  - Uniqueness of bats and their variation in Diets
    - Diversity
    - Ecological role
    - Pathogenic role
* Methods for characterizing microbiome and virome. As well as methods for linking the two  
  - Culture dependant vs independent
    - What has been discovered with these methods
    - How these methods have been applied to bats 
      - What has been found
      - What has yet to be discribed
- Goal of this study

## Materials and Methods
* Sample location discription
* Sample collection
* Sample processing
### Phase genomics portion
* (AND)

### In house analysis of MAGs, vcontigs, and viral-host pairing
#### vOTU curration
* (AND) Genomad (v x.x.x) was used to varify contigs identified as viral by phase genomics and for taxonomic identification. 
* (AND) We also ran all assembly files from phase genome through genomad to identify potential viral sequences that the phase genome pipeline may have missed
* (AND) Quality filtered viral sequences were then clustered into species-level equivalent viral operational taxanomic units (vOTUs) at 95% average nucleatide identity over 85% of the alignment fraction of the shorter sequence using the greedy centroid algorithm (anicalc.py and aniclust.py) from checkV (v x.x.x).

#### MAG curration
* (AND) MAGs from phase metagenomic were quality filtered using checkM (v x.x.x).
* (AND) Quality filtered MAGs were then dereplicated using dRep (v x.x.x) at 99% ANI with the following settings (--S_algorithm fastANI, -comp 50, --SkipMash)
* (AND) MAGs representatives were taxonomically identified using the Genome Taxonomy Database Toolkit (GTDB-Tk v x.x.x)

#### vOTU and MAG coverage
* (AND) Quality controled reads were mapped to the vOTUs and representative MAGs using bowtie (v x.x.x) and sam files were converted to indexed and sorted bam files using samtools (v x.x.x).
* (AND) Bam files were then fed into anvi'o (v x.x.x) to calculate the Q2Q3 coverage of vOTUs and MAGs across bat species.
* (AND) Q2Q3 coverages for both vOTUs and MAGs were then normalized by sequence length.
* (AND) For alpha diveristy analyzes, normalized Q2Q3 coverages for both vOTUs and MAGs were rarefied using the rrarify function of the vengan  R package.
* (AND) For beta diversity analyzes, normalized Q2Q3 coverages for both vOTUs and MAGs were used to create rarefied Hellinger distance matrices using the vegan (v x.x.x) and labdsv (v x.x.x) R packages.
* (AND) For abundance pattern analyzes, normalized Q2Q3 coverages were coverted into units of GCPM (Genome Copy Per Million Reads) for MAGs, as laid out in Rogers et al. 2022 (citation), and TPM (Transcripts Per Million Reads) for vOTUs.
* (AND) Abundance correlations of vOTUs and MAGs across and within samples were determined using hierarchical clustering of Spearman rank correlation distance matrice in base R (v x.x.x).

#### Metabolic predictions
* The program metacerberus (v x.x.x) was used for gene anotation of both viral sequences and MAGs using the following data bases: Functional Ontology Assignments for Metagenomes (FOAM), KEGG, CAZy/dbCAN, VOG, pVOG, PHROG, and COG.

#### AMR and CAZY gene abundance patterns
* (AND) Read coverage and abundance for AMR and CAZY genes were calculated using a custom ihhouse script.

#### Linking virus to host
* (AND) IPHoP (v x.x.x) was used to link all viral sequences to all medium to high quality MAGs, not just to the representative vOTUs and MAGs. 
* (AND) The viral and MAG IDs of those within the viral-host predictions from both IPHoP and Phase genomes were replaced with the corresponding representative vOTU and MAG IDs.

#### Statistical analysis
* (AND) Differences in MAG and vOTU composition among bat species were analyzed via distance based redundancy analysis (db-RDA) on a quntitative Hellinger distance matirx. 
* (AND) The richness and evenness of MAGs and vOTUs were compared using the R package microbiome for richness (chao1) and the package hillR (v x.x.x) for evenness (Pielou's index).
* (AND) Statistical significants was determined based on 9999 permutations of the data in the vegan R package.

## Results
* MAG
  - Number and quality of MAGs
  - Taxonomic idenification
  - Distribution patterns
    - Alpha Diversity
      - Richness
      - Evenness
    - Beta Diversity
    - Metabolic profiles of gut microbiome across bat species

* vOTUs
  - Number of vOTUs
  - Quality of vOTUs
  - Taxonomic idenification
  - Distribution patterns
    - Alpha Diversity 
      - Richness
      - Evenness
    - Beta Diversity
    - AMGs and viral vectors

* Virus-host matching
  - IPHop
    - CRISPR
    - tetranucleotide frequency
    - blast
  - HiC

### MAGs
```{r, echo=FLASE}
# Quality and length of rarefied MAGs
mag_stats<-fread("/projects/raw_lab/projects/Bats/bat_gut_micro/phasegenome_out/data/results/tables/mag_stats.csv")
# Abudance information:
mag_fam_rel_abun<-fread("/projects/raw_lab/projects/Bats/bat_gut_micro/phasegenome_out/data/results/tables/mag_fam_rel_abun.csv")
phylum_rel_abun<-mag_fam_rel_abun%>%select(-V1,-class:-family)%>%
  aggregate(.~phylum,sum)
colSums(phylum_rel_abun[,-1])
```
* (AND) We recovered `r mag_stats[domain=="Archaea",.N]` Archaeal and `r mag_stats[mag_stats$domain=="Bacteria",.N]` Bacterial medium to high quality MAGs with $\geq$ `r paste0(round(min(mag_stats$completeness)),"%")` and $\lt$ `r paste0(round(max(mag_stats$contamination)),"%")` redundancy.
* (AND) Of these `r nrow(mag_stats)` MAGs, `r length(mag_stats$genome[mag_stats$completeness>=90 & mag_stats$contamination<10,])` were $\geq$ 90% complete and $\lt$ 10% redundant, while YY are of high quality as determined by parameters layed out in XXX et al. (** TABLE **).
* (AND) Bacillota, Actinomycetota, and Pseudomonadata were the predominate phylum across the gut microbiome within all 3 of the bat species.
* (AND) No significant differnces were found in the richness (Choa1) or evenness (Pielou's Index) of prokaryotic MAGs across the 3 bat species.
* (AND) db-RDA analysis revealed no significant differnce in the $B$-diversity of the prokaryotic MAGs across the 3 bat species.
\includegraphics{../data/results/figures/FigX_top10_most_abun_viral_taxa.pdf}
### vOTUs
```{r,echo=FALSE}

```
- A total of XX vOTUs were recovered.
- X, Y, and Z were the most predominate viral taxonomic unit across the gut virome with all 3 of the bat species.

#### Metabolic capabilities
* (AND)
## Descusion

## Conclusion

## Acknowledgement

### 2025-03-03 snakemake_project/workflow/scripts/host_read_removal.sh
Updating repository

@@ -1,45 +1,0 @@
-#!/usr/bin/env bash
-
-################################################################
-# author: TJ Rogers                                            #
-# input: Name of the trimmed files (without the path)          #
-# output: places cleaned reads into data/processed/clean_reads #
-#                                                              #
-# Removing host sequences.                                     #
-################################################################
-
-#Assigning arguments:
-sample=$(basename ${$1} -sg_1.fastq.gz) #Grabing the sample name
-r1=$1                                   #Forward read input
-r2=$2                                   #Reverse read input
-fr1=$3                                  #Forward read output
-fr2=$4                                  #Reverse read output
-host_genome=$5                          #Host genome to index
-
-#### Create bowtie2 database of the host genome
-bowtie2-build "$host_genome" host_DB
-
-#### Map reads against host sequence database, keeping both mapped and unmapped reads
-bowtie2 -p 8 -x host_DB \
--1 "$r1" \
--2 "$r2" \
--S "$sample"_mapped_and_unmapped.sam
-
-#### Convert sam files to bam files to save space
-samtools view -bS "$sample"_mapped_and_unmapped.sam > "$sample"_mapped_and_unmapped.bam
-
-#### Collect reads not mapped to host sequences
-samtools view -b -f 12 -F 256 \
-    "$sample"_mapped_and_unmapped.bam \
-    > "$sample"_bothReadsUnmapped.bam 
-
-#### Sort bam files of unmapped reads
-samtools sort -n -m 5G -@ 2 "$sample"_bothReadsUnmapped.bam -o "$sample"_bothReadsUnmapped_sorted.bam
-
-#### Separate the paired-end reads into separate fastq files
-samtools fastq -@ 8 "$sample"_bothReadsUnmapped_sorted.bam \
-  -1 "$fr1" \
-  -2 "$fr2" \
-  -0 /dev/null -s /dev/null -n
-
-#### Cleaning up tmp files

### 2025-03-03 snakemake_project/workflow/scripts/retreve_host_references.sh

@@ -1,39 +1,0 @@
-#!/usr/bin/env bash
-
-################################################################
-# author: TJ Rogers                                            #
-# input: Name of the clean files (without the path)            #
-# output: places filterd cleaned reads into                    #
-# data/processed/clean_reads                                   #
-#                                                              #
-# Removing host reads from clean reads.                        #
-################################################################
-
-
-###########################################################
-# This code block dowloads the genomes for representative #
-# sequences using the datasets program.                   #
-###########################################################
-#### Calling in input and output from the rule doc:
-accession=$1
-ncbi_zip=$2
-ncbi_unziped=$3
-refdir=$4
-#### Running datasets to retrieve sequences
-datasets download genome accession --inputfile "$accession" \
-    --include gff3,rna,cds,protein,genome,seq-report \
-    --filename "$ncbi_zip"
-
-unzip "$ncbi_zip" -d "$refdir"
-
-mv "$refdir"ncbi_dataset/data/GCA_036768555.1/GCA_036768555.1_ASM3676855v1_genomic.fna "$refdir"ncbi_dataset/data/GCA_036768555.1/GCA_036768555.1.fna
-mv "$refdir"ncbi_dataset/data/GCA_028538775.1/GCA_028538775.1_mMyoYum1.0.hap1_genomic.fna "$refdir"ncbi_dataset/data/GCA_028538775.1/GCA_028538775.1.fna
-mv "$refdir"/ncbi_dataset/data/GCA_004126475.3/GCF_004126475.2_mPhyDis1.pri.v3_genomic.fna "$refdir"ncbi_dataset/data/GCA_004126475.3/GCF_004126475.3.fna
-mv "$refdir"/ncbi_dataset/data/GCA_019186645.2/GCF_019186645.2_TTU_PhHast_1.1_genomic.fna "$refdir"ncbi_dataset/data/GCA_019186645.2/GCF_019186645.2.fna
-##########################################################
-# This code chunk was originally used to test the script #
-##########################################################
-# datasets download genome accession --inputfile "$refdir"/test_bat_host_accessions.txt \
-#     --include gff3,rna,cds,protein,genome,seq-report --filename /projects/raw_lab/projects/Bats/bat_gut_micro/test/bat_references.zip
-
-# unzip /projects/raw_lab/projects/Bats/bat_gut_micro/test/bat_references.zip /projects/raw_lab/projects/Bats/bat_gut_micro/test/

### 2025-03-03 snakemake_project/workflow/scripts/trimming.sh

@@ -1,39 +1,0 @@
-#!/usr/bin/env bash
-
-################################################################
-# author: TJ Rogers                                            #
-# input: Name of the raw files (without the path)              #
-# output: places trimed reads into data/temp                   #
-#                                                              #
-# Trimming the raw reads.                                      #
-################################################################
-
-###################################################
-# Send all stderr from this script to the log file #
-###################################################
-
-#####################################
-# Call read files from the rule doc #
-#####################################
-
-#####################
-# Removing adapters #
-#####################
-# Check if final output files exist
-if [[ -f "$5" && -f "$6" ]]; then
-  echo "Final files $5 and $6 already exist. Skipping this sample..."
-  exit 0
-fi
-
-# Check if intermediate files exist
-if [[ -f "$3" && -f "$4" ]]; then
-    echo "Intermediate files $3 and $4 found. Running step 2 to remove phiX sequences..."
-    # Removing phiX
-    bbduk.sh -Xmx1g in1=$1 in2=$2 out1=$3 out2=$4 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/adapters.fa ktrim=r k=21 qin=auto mink=11 hdist=2 tbo tpe
-else
-    echo "Intermediate files not found. Running both steps to remove adapter and phiX sequences..."
-    # Removing adapters
-    bbduk.sh -Xmx1g in1=$1 in2=$2 out1=$3 out2=$4 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/adapters.fa ktrim=r k=21 qin=auto mink=11 hdist=2 tbo tpe
-    # Removing phiX
-    bbduk.sh -Xmx1g in1=$3 in2=$4 out1=$5 out2=$6 ref=/users/troger50/.conda/envs/bbmap/share/bbmap/resources/phix174_ill.ref.fa.gz k=27 hdist=1 qtrim=rl trimq=17 cardinality=t mingc=0.05 maxgc=0.95
-fi

### 2025-03-03 titan_project/bin/scripts/titan_binning2.py

@@ -1,214 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import subprocess
-import argparse
-from glob import glob
-
-###########################################
-# Make sure everything is set up properly #
-###########################################
-# The default length (-l) setting is 1000bp.
-l=1000
-
-#########################################
-# Aligning reads to make coverage files #
-#########################################
-# Coverage file(s) for the binning programs
-# I need to create coverage files for each binning program. This will be coverage
-# files of each assembly accross all metagenomes. This will allow for better bin-
-# ing.
-def align_reads(read_path, assembly, output_dir):
-    os.makedirs(output_dir, exist_ok=True)
-
-    # Find forward and reverse reads
-    R1 = sorted(glob(os.path.join(read_path, '**', "*_R1_001.fastq.gz"), recursive=True))
-    R2 = sorted(glob(os.path.join(read_path, '**', "*_R2_001.fastq.gz"), recursive=True))
-
-    # Check that forward and reverse basenames match
-    if sorted(os.path.basename(f).replace('_R1_001.fastq.gz', '') for f in R1) != \
-       sorted(os.path.basename(r).replace('_R2_001.fastq.gz', '') for r in R2):
-        raise ValueError("Mismatch between forward and reverse reads. Ensure that each forward read has a matching reverse read.")
-    
-    reads = list(zip(R1, R2))
-    print(f"Indexing assembly: {assembly}")
-    assembly_name = os.path.basename(os.path.dirname(os.path.dirname(assembly)))
-    assembly_output_dir = os.path.join(output_dir, assembly_name)
-    os.makedirs(assembly_output_dir, exist_ok=True)
-    ref_index= os.path.splitext(os.path.basename(assembly))[0]
-    ref_index_path=os.path.join(assembly_output_dir, f"{ref_index}")
-    subprocess.run(f"mamba run -n bwa bwa index -p {ref_index_path} {assembly}", shell=True, check=True)
-
-
-    for i, (forward, reverse) in enumerate(reads):
-        sample_name = "_".join(os.path.basename(forward).split("_")[:-2])
-        sam_output = os.path.join(assembly_output_dir, f"{sample_name}.sam")
-        unbam_output = os.path.join(assembly_output_dir, f"{sample_name}.unsorted.bam")
-        bam_output = os.path.join(assembly_output_dir, f"{sample_name}.sorted.bam")
-        # Align reads with BWA and convert to BAM using samtools
[…]

### 2025-06-30 phasegenome_out/doc/paper/draft_manuscript.rmd
Manuscript update

---
title: "Resolving gut microbiome networks within Chiropterans"
author: "Timothy J. Rogers, Richard A. White, Laurel R. Yohe†"
affiliation: "Department of Bioinformatics, University of North Carolina, Charlotte, NC; Department of Bioinformatics and Genomics, North Carolina Research Center (NCRC), Kannapolis, NC"
output:
  word_document:
    reference_docx: ~/Documents/draft_rmd_styles/word-styles-reference-01.docx
---

${\dagger}$ To whom correspondence should be addressed

\newpage
```{r, include=FALSE}
library(data.table)
library(tidyverse)
library(phyloseq)
```

```{r,include=FALSE}
# Useful functions:
join_with_and <- function(x) {
  if(length(x) > 1){
    paste(paste(x[-length(x)], collapse = ", "), "and", x[length(x)])
  } else {
    x
  }
}
```
# Abstract

\newpage

# Introduction
* Bats are known carriers of human associated pathogens
* The reason bats are bioreactors is not understood
* The diet of bats may contribute to the gut microbiota makeup
* Phage associated with these microbiota can benifit and hinder microbial populations and have an impack on bat immune responses
* Some ideas suggest viral tolerance is linked to 
  - Uniqueness of bats and their variation in Diets
    - Diversity
    - Ecological role
    - Pathogenic role
* Methods for characterizing microbiome and virome. As well as methods for linking the two  
  - Culture dependant vs independent
    - What has been discovered with these methods
    - How these methods have been applied to bats 
      - What has been found
      - What has yet to be discribed
- Goal of this study

```{r FigX, echo = FALSE, message=FALSE, fig.cap='**Figure X. Bat Holobiont**'}

  knitr::include_graphics("../../data/results/figures/")
```
# Materials and Methods
* Sample location discription
* Sample collection
* Sample processing
### Phase genomics portion
* (AND)

## Sample collection
### ProxiMeta Hi-C and metagenomic pipeline
  Guano samples were then used to generate Hi-C libraries using the protocol included in the ProxiMeta Hi-C kit (**Phase Genomics**). Hi-C operates by chemically cross-linking DNA fragments that lie in close spatial proximity inside intact cells, allowing for active phage infection identification. High quality Hi-C libraries were produced for **X** guano samples from **X** bat species. **X** libraries were successfully extracted and sequenced from bat species **X**, **Y**, and **Z**. Briefly, extracted genomic DNA along with chemically fixed fecal samples were sent to Phase Genomics (Seattle, WA, USA) for Hi-C proximity ligation and sequencing. Hi-C libraries were constructed using the Phase Genomics ProxiMeta Hi-C v4.0 Kit, while metagenomic shotgun libraries were generated in parallel with ProxiMeta’s standard library preparation reagents. Both library types were sequenced on an Illumina NovaSeq instrument to produce 150 bp paired-end reads. (**Need to add in here the methods that ProxiMeta uses to construct the MAGs**). The resulting **X** million Hi-C and **X** million shotgun metagenomic paired-end read fastq files were then uploaded to the Phase Genomics cloud portal for downstream analysis.

## Microbial Community Profiling
  SingleM (vXXX; **CITATION**) "pipe" was used to extract metagenome operational taxonomic units (mOTUs) from 59 single copy conserved genes (22 Bacteria-specific, 24 Archaea-specific, and another 13 targeting both domains) identified in the raw reads and create an OTU table. Then, SingleM "pipe" was ran on the dereplicated metagenome assembled genomes (MAGs; see below) as well as the unbinned contigs. The OTU tables from the raw reads, MAGs, and unbinned contigs were then compared using SingleM "appraise" using a sequence identity of 89% to determine how much of the metagenomes were represented in the assemblies and MAGs at the genus level.

### MAG curration and taxonomic identification
  The ProxiMeta produced a total of 441 medium to high quality MAGs ($\geq$ 50% complete with $\leq$ 10% redundancy) across all bat species. MAGs were retrieved from the Phase Genomics cloud portal and stored on the University of North Carolina at Charlotte's high-performance computing (HPC) cluster. To allow for comparision of MAGs across samples, dRep (**CITATION**) was used to dereplicate the MAGs at 98% ANI using the command “dRep cluster --SkipMash --S_algorithm fastANI -comp 50”. This resulted in a final MAG set of 239 medium to high quality dereplicated MAGs. MAG quality was determined by CheckM2 (**v x.x.x**). MAG taxonomy was assigned using the Genome Taxonomy Database Toolkit (GTDB-Tk(**CITATION**)). GTDB nomenclature is used throughout the manuscript. However, other commonly used nomenclatures are used when they aid in clarity.

### vOTU curration
  Viral contigs (vContigs) identified by ProxiMeta were retrieved from the Phase Genomics cloud portal. To identify potential vContigs not identified by ProxiMeta, genomade (**V**) was used to scan the assembly files. Anvi'o (**V**) anvi-script-reformate was using to clean up all identied vContig headers for downstream analysis. From there, all vContigs were combined into a single fasta file. They were then clustered into viral operational taxonomic units (vOTUs) at 95% average nucleotide identity (ANI) over 85% of the alignment fraction of the shorter sequence using the greedy centroid algorithm (anicalc.py and aniclust.py) from CheckV (v x.x.x). vOTUs were then ran through CheckV for quality scores and lifestyle predictions. In addition to CheckV, BACPHLIP (**V**) was also used in addition to finer scale resolution of viral lifestyle of the vOTUs. Viral lifestyle (lytic or lysogenic) was determined by integrating predictions from CheckV and BACPHLIP. The CheckV "provirus" annotation was considered, where a value of "Yes" indicated the presence of a provirus and suggested a lysogenic lifestyle. BACPHLIP outputs included probabilities for "Virulent" and "Temperate" lifestyles, each ranging from 0 to 1. A vOTU was classified as lysogenic if CheckV annotated it as a provirus ("Yes"), or if the BACPHLIP "Temperate" probability was ≥0.5. A contig was classified as lytic if CheckV did not detect a provirus ("No") and the BACPHLIP "Virulent" probability was ≥0.75, or ≥0.5 if the genome was estimated to be 100% complete. If neither tool confidently predicted the lifestyle, the contig was labeled as "unclear." The method used for each final prediction was also recorded, indicating whether the lifestyle assignment was based on CheckV, BACPHLIP, or a combination of both.

### vOTU and MAG coverage
  Quantification of dereplicatied MAGs and vOTUs across all samples was calculated using the anvi'o tool. Briefly, quality controlled reads were mapped to the vOTUs and representative MAGs using bowtie (v x.x.x) and sam files were converted to indexed and sorted bam files using samtools (v x.x.x) (**CITATION**). The resulting BAM files for the MAGs and vOTUs were then used by the anvi'o tool kit to calculate the  depth of coverage across nucleotide positions in which coverage was within the interquartile range (Q2Q3). Q2Q3 coverage was used to filter out outliers in coverage caused highly similar sequences shared among different taxa. MAG abundance was quantified in units of weight mean transcripts per million reads (wtTPM) for contigs within each MAG. First, we calculated the transcripts per million reads (TPM) for each contig at each site (**equation 1**).

  **Equation 1:**
  $\mathrm{TPM}_i = \frac{c_i/L_i}{\sum_{j=1}^n c_j/L_j} \times 10^6$

  In this equation, $TPM_i$ denotes the Transcripts Per Million value for contig ${i}$, representing the normalized abundance of that contig after adjusting for both contig length and total sequencing depth. The variable ${c}_i$ refers to the number of sequencing reads mapped to contig ${i}$, while ${L}_i$ is the length of contig ${i}$ in base pairs. We then multiply by one million ($10^6$) to convert the values into TPM units.To then quantify MAG abundance across samples, we then calculated the wtTPM using the below equation 2

 **Equation 2:**
 $\mathrm{wtTPM}_m = \frac{\sum_{c=1}^n TPM_i \times L_i}{\sum_{c=1}^n L_i}$

In this expression, $wtTPM_m$ denotes the length-weighted mean TPM for the total contigs within metagenome-assembled genome (MAG) $m$. vOTU abundance was quantified as transcripts per million reads (TPM; Equation 1).

### vOTU and MAG metabolic predictions
  Metacerberus (**CITATION**) was used for open reading frame prediction, gene annotation, KEGG orthology identifier linkage, and metabolic predictions using the Functional Ontology Assignments for Metagenomes (FOAM; (**CITATION**)), KEGG (**CITATION**), CAZy/dbCAN (**CITATION**), VOG (**CITATION**), pVOG (**CITATION**), PHROG (**CITATION**), and COG (**CITATION**) databases. An in house script was usied to identify AMR and CAZY genes as well as calculate their coverage/abundance.

### Linking virus to host
  In addition to the host associations predicted through proximity ligation by ProxiMeta, iPHoP (Integrated Phage HOst Prediction; **Roux et al. 2023**) was also used to identify additional potential host associations. iPHoP uses six host prediction methods to determine phage-host interactions (typically at a genus or family rank): Blast, CRISPR spacer matching  (**CITATION**), three different oligonucleotide frequency (ONF) comparison methods (VirHostMatcher (**CITATION**), WIsH (**CITATION**), and PHP (**CITATION**)), and Random Forest Assignment of Hosts (RaFAH; (**CITATION**)). First, iPHoP was ran with default parameters to predict host of a vOTUs based on iPHoP's default database (**DATA BASE VERSION**). Then, the original 441 MAGs from ProxiMeta were added to the default database using (using the "add_to_db" function) to create a custom database. This custom database was used to run a final host prediction on the same vOTUs. Finally, phage-host pairings were filtered using a minimum cutoff pairing score of 75. Only host paring predictions to MAGs were considered for down stream analysis. If a vOTU was paired to a MAG that was not a representative MAG via proximity ligation or one of iPHoPs methods, then that pairing was given to the representative MAG for downstream analysis.

### Statistical analysis
  Results were imported into R for analysis and visualization using the R packages data.table (**CITATION**), tidyverse (**CITATION**), ggplot (**CITATION**), vegan (**CITATION**), hillR (**CITATION**), phyloseq (**CITATION**), ggtext (**CITATION**), and ggpubr (**CITATION**). mOTUs identified by SingleM were used to assess alpha and beta diversity. Differences in prokaryotic community composition among bat species were analyzed using distance-based redundancy analysis (db-RDA) on a Hellinger distance matrix. Statistical significance was evaluated with 9,999 permutations using the R package vegan (**V**) (**Oksanen et al. 2015**). Richness (Chao1), evenness (Pielou's index), and alpha diversity (Hill numbers at q = 1, corresponding to the inverse Shannon index) were calculated. Comparisons of alpha diversity across bat species were performed using two-way ANOVA.
  To identify microbial mOTUs associated with each bat species, we performed an indicator species analysis following the approach of **De Cáceres and Legendre (2009)** and **De Cáceres et al. (2010)**. Group-equalized correlation coefficients (r₉) were calculated for each bat species to account for differences in taxon abundance and enable direct comparisons among taxa with varying relative abundances. Statistical significance was assessed through 9,999 permutations. All analyses were carried out using the R package indicspecies (**V**) (**De Cáceres et al. 2016**). Subsequently, the relative abundances of indicator mOTUs were calculated for each bat species.

# Results
## SingleM analysis
```{r,include=FALSE}
# Before rarifaction
single_otu<-fread("../../data/results/tables/singlem_taxa.csv")

# After rarifaction
#Calling in phyloseq object
rel_abun_phy<-readRDS("../../data/results/tables/singleM__rarefied_phy.rds")

```
  Bat gut microbiomes were profiled based on metagenome operational taxonomic units (mOTU) using `r length(unique(single_otu$marker))` single copy conserved genes identified by SingleM. After rarefaction, and removal of Eukaryota sequences, a total of `r nrow(otu_table(rel_abun_phy))` mOTUs were retained. tb-RDA on the Hellinger distance matrix revealed significant differences in the gut prokaryotic community composition across the three bat species (**Fig. 1**, _adj. R^2^_=0.27, _F_=2.3, _P_ < 0.01). However, pairwise comparisons were unable to detect differences across the bat species.

```{r Fig1, echo = FALSE, message=FALSE, fig.cap='**Figure 1. Transformation based redundancy analysis of the bat gut prokaryotic community mOTUs.**'}

  knitr::include_graphics("../../data/results/figures/new_singlem_dbrda.pdf")
```

```{r,include=FALSE}
#Calculating relative abundance of phylum
phy_per<-((table(tax_table(rel_abun_phy)[,"phylum"])/sum(table(tax_table(rel_abun_phy)[,"phylum"])))*100)%>%
  sort(.,decreasing=TRUE)%>%data.frame%>%
  rename(phylum=Var1)%>%
  mutate(phy_percent=ifelse(Freq==max(Freq),paste0(round(Freq),"% of the mOTUs belonged to _",phylum,"_"),paste0(round(Freq),"% belonged to _",phylum,"_")))

#Calling in indicator analysis csv
ind_sp<-fread("../../data/results/tables/median_final_indicator_table.csv")%>%
  mutate(across(`Mimon crenulatum`:`Phyllostomus hastatus`, ~as.numeric(gsub("\\*\\*", "", .))))

# Indicator species analysis results for singlem
# sm_indic<-fread("../../data/results/tables/bat_gut_singlem_indicator_sp.csv")

```

### Prokaryotic community differences across the gut of bat species
  Across bat species, `r paste0(phy_per$phy_percent[phy_per$Freq==max(phy_per$Freq)],", ", join_with_and(phy_per$phy_percent[phy_per$Freq>1&phy_per$Freq<max(phy_per$Freq)]))`. The remaining phylum constituted < 1% each. Prokaryotic $\alpha$-diversity (inverse-Shannon) did not significantly differ across bat species. However, Prokaryotic community evenness was significantly lower in _Phyllostomus discolor_ than in _P. hastus_ (**Fig. 2C**, P<0.05). Even so, no significant difference was found in the prokaryotic richness (Chao1). A total of `r length(unique(ind_sp$OTU_ID))`Indicator species analysis, based on r~g~, identified specific mOTUs common in some bat species, but in lower relative abundance in others (**Table SX**).

  For example, members of the bacteria pylum _Bacillota_, _Loactococcus lactis_ and unidentified taxa in the genera _Loactococcus_ and  _Streptococcus_, had high relative abundances in _Mimon crenulatum_ (10%, 10%, and 6% respectively) while they were not found at all in _P. discolor_ nor _P. hastus_. Unidentified taxa in the bacteria genus _Arachnia_ and another in the domain _Archean_ were abundant in _P. discolor_ (.03% and .01% respectively) but were not present in _P. hastus_ nor _M. crenulatum_. Conversely, members of the bacteria _Bacillota_, an unidentified taxa in the genus _Gemella_ and the species _Gemella sanguinis_, had higher abundances in _P. hastus_ (35%, and 3% respectively) when compared to _M. crenulatum_ or _P. discolor_.

```{r Fig2, echo = FALSE, message=FALSE, fig.cap='**Figure 2. Bat gut prokaryotic community diversity of the mOTUs.**', out.height = "800px", out.width='480px'}

  knitr::include_graphics("../../data/results/figures/singleM_rarefied_a-diversity.pdf")
```

## Metagenomic analysis

### MAG diversity and community representation
```{r, include=FALSE}
# Quality and length of rarefied MAGs
mag_stats<-fread("../../data/results/tables/MAG_quality.csv")
# Abudance information:
#mag_fam_rel_abun<-fread("../../data/results/tables/mag_fam_rel_abun.csv")
#phylum_rel_abun<-mag_fam_rel_abun%>%select(-V1,-class:-family)%>%
  #aggregate(.~phylum,sum)
#colSums(phylum_rel_abun[,-1])
```
  Assembly, binning, and dereplication at the species level resulted in the recovered of `r mag_stats[domain=="Archaea",.N]` Archaeal and `r mag_stats[mag_stats$domain=="Bacteria",.N]` Bacterial medium to high quality MAGs with $\geq$ `r paste0(round(min(mag_stats$completeness))," %")` and $\leq$ `r paste0(round(max(mag_stats$contamination))," %")` redundancy. Of these `r nrow(mag_stats)` MAGs, `r length(mag_stats$MAG[mag_stats$completeness>=90 & mag_stats$contamination<10])` were $\geq$ 90% complete and $\leq$ 10% redundant, while **X** meet the MIMAG standard of high quality draft genomes ((**CITATION**), **Table**). **X%** (n=**X**) of the MAGs were prepredicted to be novel by GTDB-Tk (**what tax levels**).

  These MAGs represented between 16% and 78% of the whole community, with a median of 61.4% based on the genus-level recovery estimentes ((**Singlem Citation**), **TABLE X**). _Bacillota_, _Actinomycetota_, and _Pseudomonadata_ were the predominate phylum across the gut microbiome MAGs within all 3 of the bat species (**Fig 3**).

```{r Fig3, echo = FALSE, message=FALSE, fig.cap='**Figure 3. Top 10 most abundant MAG phylum from across the bat gut metagenomes**', out.height = "700px", out.width='800px'}

knitr::include_graphics("../../data/results/figures/FigX_top10_most_abun_mag_taxa.pdf")
```

### vOTUs
```{r,include=FALSE}
vOTU_rar_cov<-fread("../../data/results/tables/vOTU_rarified_counts.csv")
vOTU_rel_abun<-fread("../../data/results/tables/viral_rel_abun.csv")
vOTU_quality<-fread("../../data/results/tables/quality_summary.tsv") %>%
  filter(contig_id %in% colnames(vOTU_rar_cov[,-1]))

```
  A total of **17702** potential viral sequences were ID from the 6 individual assemblies. Clustering at 95% ANI over 85% of the shortest sequence identified 16289 viral operational taxanomic units (vOTUs) that were $\geq$ 1 kbp in length. Further filtering for sequences $\geq$ 2.5 kpb resulted in a final set of 6235 vOTUs. After rarefaction, `r ncol(vOTU_rar_cov[,-1])` vOTUs were retained. CheckV was used to assess the quality of these sequences, revealing that `r nrow(vOTU_quality[vOTU_quality$completeness>=45.5])` (`r round(nrow(vOTU_quality[vOTU_quality$completeness>=45.5])/nrow(vOTU_quality)*100)`%) were $\geq$ 50% complete including `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Complete"])` complete vOTUs that identified on the bases of direct terminal repeats (DTR), `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5])` high quality vOTUs that were identified on the bases of AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="High-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs), `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5])` medium quality vOTUs that were idenified on the bases on AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI-based",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Medium-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs), and `r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5])` low quality vOTUs that were identified based on AAI (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5 & grepl("AAI-based",vOTU_quality$completeness_method)])` vOTUs) and HMM (`r nrow(vOTU_quality[vOTU_quality$CheckV_quality=="Low-quality" & vOTU_quality$completeness>=45.5 & grepl("HMM",vOTU_quality$completeness_method)])` vOTUs). The reset of the vOTUs (`r nrow(vOTU_quality[vOTU_quality$completeness<45.5|is.na(vOTU_quality$completeness)])`) were of low quality (`r nrow(vOTU_quality[vOTU_quality$completeness<45.5])`) or the quality was undetermined (`r nrow(vOTU_quality[is.na(vOTU_quality$completeness)])`). An unclassified order of the class _Caudoviricetes_, families _Retroviridae_, _Adintoviridae_, and _Iridoviridae_, an unclassified family of _Kyanoviridae_, families _Inoviridae_ and _Bornaviridae_, an unclassified family of _Herelleviridae_, families _Mimiviridae_ and _Parvoviridae_ were the most predominate viral taxa across the gut virome with all 3 of the bat species (**Figure 4**). No statistically significant differences were found in the viral richness, envenness, $\alpha$-diversity, or $B$-diversity nor did indicator analysis reveal any indicator viral spices for the bat species.

```{r Fig4, echo = FALSE, message=FALSE, fig.cap='**Figure 4. Top 10 most abundant viral taxa from across the bat gut metagenomes**', out.height = "700px", out.width='800px'}

knitr::include_graphics("../../data/results/figures/FigX_top10_most_abun_viral_taxa.pdf")
```

### Virus-host predicitons
```{r,include=FALSE}
# Iphope and phase virus-host predictions #
###########################################
ip_ph_compare<-fread("../../data/results/tables/iphop_phase_compare.csv")
ip_ph_compare_w_indicators<-fread("../../data/results/tables/ip_ph_compare_with_indicator_info.csv")

# IPhop and phase virus-host predictions at the taxonomic level #
#################################################################
ip_ph_compare_taxa<-fread("../../data/results/tables/iphop_phase_compare_taxa_level.csv")

# Host Phylum target frequency #
################################
# Compute the aggregate and filter
df_filtered <- unique(ip_ph_compare[, c("MAG", "phylum")]) %>% 
  mutate(num_mags=1)%>%
  aggregate(num_mags ~ phylum, sum) %>%
  arrange(desc(num_mags)) %>%
  filter(num_mags >= 10)
# Wrap each phylum with asterisks for italics in Markdown
df_filtered$phylum <- paste0("*", df_filtered$phylum, "*")
# Create nicely formatted strings for phyla and their corresponding num_mags
join_with_and <- function(x) {
  if(length(x) > 1){
    paste(paste(x[-length(x)], collapse = ", "), "and", x[length(x)])
  } else {
    x
  }
}

phy_string <- join_with_and(df_filtered$phylum)
mags_string <- join_with_and(df_filtered$num_mags)

host_fam_count<-unique(ip_ph_compare_taxa[, c("phylum", "host_family","num_mags")])%>%
  arrange(desc(num_mags)) %>% 
  mutate(phylum=paste0("*", phylum, "*"),
         host_family=paste0("*", host_family, "*"))%>% 
  filter(phylum %in%df_filtered$phylum)

fam_string_join<-function(df,phy){
  phy_select<-df%>%
    filter(phylum==phy)%>%
    { if (any(.$num_mags >= 5)) {
      filter(., num_mags >= 5)
    } else {
      filter(., num_mags == max(num_mags))
    }
  }
  fam_join<-join_with_and(phy_select$host_family)
  num_join<-join_with_and(phy_select$num_mags)
  return(list(fam_join,num_join))
}
# Table of phage-host pairing
table_df <- as.data.frame(table(ip_ph_compare$lifestyle, ip_ph_compare$iphop, ip_ph_compare$phase))%>%
  subset(Freq!=0)

names(table_df) <- c("lifestyle", "iphop", "phase", "Freq")

table_df<-table_df%>%
  mutate(prediction=case_when(iphop==TRUE&phase==TRUE~"both",
                              iphop==TRUE&phase==FALSE~"iphop",
                              iphop==FALSE&phase==TRUE~"phase"))%>%
  select(-iphop,-phase)%>%
  group_by(prediction) %>%
  mutate(percent_within_group = round(Freq / sum(Freq) * 100)) %>%
  ungroup()
```

  Phage-host detected by Hi-C were compared to those made by the bioinformatics pipeline iPHoP. Using both Hi-C and iPHoP methods, a total of `r nrow(unique(ip_ph_compare[,c("MAG","vOTU")]))` phage-host predictions were made using phase HiC and IPHoP methods. These predictions included  `r round(length(unique(ip_ph_compare$MAG))/nrow(mag_stats)*100)`% (n= `r length(unique(ip_ph_compare$MAG))`) of the total MAGs and **x/y*100**% (n=`r length(unique(ip_ph_compare$vOTU))`) of the total vOTUs.

  Hi-C analysis allowed for the identification of `r nrow(ip_ph_compare[iphop==FALSE&phase==TRUE])` unique phage-host links comprised of `r length(unique(ip_ph_compare[ip_ph_compare$phase=="TRUE"&ip_ph_compare$iphop=="FALSE"]$MAG))` MAGs and `r length(unique(ip_ph_compare[ip_ph_compare$phase=="TRUE"&ip_ph_compare$iphop=="FALSE"]$vOTU))` vOTUs. Using the six methods employed by iPHoP, a total of `r nrow(ip_ph_compare[iphop==TRUE&phase==FALSE])` unique phage-host links, comprised of `r length(unique(ip_ph_compare[ip_ph_compare$phase=="FALSE"&ip_ph_compare$iphop=="TRUE"]$MAG))` MAGs and `r length(unique(ip_ph_compare[ip_ph_compare$phase=="FALSE"&ip_ph_compare$iphop=="TRUE"]$vOTU))` vOTUs, were identified. We also identified `r nrow(ip_ph_compare[iphop==TRUE&phase==TRUE])` unique phage-host links that were identified by both methods (**Figure 5**). These were comprised of `r length(unique(ip_ph_compare[ip_ph_compare$phase=="TRUE"&ip_ph_compare$iphop=="TRUE"]$MAG))` MAGs and `r length(unique(ip_ph_compare[ip_ph_compare$phase=="TRUE"&ip_ph_compare$iphop=="TRUE"]$vOTU))` vOTUs. Of those predictions made specifically by IPHoP, `r nrow(ip_ph_compare[iphop==TRUE&phase==FALSE&method_iphop=="blast"])` are based on blastn and `r nrow(ip_ph_compare[iphop==TRUE&phase==FALSE&method_iphop=="iPHoP-RF"])` are based on a combination of the 6 methods that iPHoP uses (called iPHoP-RF).

```{r Fig5, echo = FALSE, message=FALSE, fig.cap='**Figure 5. Venn diagram of phage-host predictions made with the Hi-C and iPHoP methods**', out.height = "800px", out.width='800px'}

knitr::include_graphics("../../data/results/figures/updated_venndiagram_phase_iphop.pdf")
```

  Of the `r nrow(ip_ph_compare[iphop==FALSE&phase==TRUE])` phage-host predictions made by phase, `r table_df[table_df$prediction == "phase"&table_df$lifestyle=="lysogenic", ]$percent_within_group`% of the pairings were with lysogenic phage, `r table_df[table_df$prediction == "phase"&table_df$lifestyle=="lytic", ]$percent_within_group`% were with lytic phage, and `r table_df[table_df$prediction == "phase"&table_df$lifestyle=="unclear", ]$percent_within_group`% were with phages whose lifestyle was not able to be predicted. Of the `r nrow(ip_ph_compare[iphop==TRUE&phase==FALSE])` phage-host predictions made by iphop, `r table_df[table_df$prediction == "iphop"&table_df$lifestyle=="lysogenic", ]$percent_within_group`% of the pairings were with lysogenic phage, `r table_df[table_df$prediction == "iphop"&table_df$lifestyle=="lytic", ]$percent_within_group`% were with lytic phage, and `r table_df[table_df$prediction == "iphop"&table_df$lifestyle=="unclear", ]$percent_within_group`% were with phages whose lifestyle was not able to be predicted. Of the `r nrow(ip_ph_compare[iphop==TRUE&phase==TRUE])` phage-host predictions made by both phase and iphop, `r table_df[table_df$prediction == "both"&table_df$lifestyle=="lysogenic", ]$percent_within_group`% of the pairings were with lysogenic phage, `r table_df[table_df$prediction == "both"&table_df$lifestyle=="lytic", ]$percent_within_group`% were with lytic phage, and `r table_df[table_df$prediction == "both"&table_df$lifestyle=="unclear", ]$percent_within_group`% were with phages whose lifestyle was not able to be predicted (**Fig. 6**).

  `r paste0("MAGs within the bacteria phyla ", phy_string, " had the highest frequency of being targeted by viruses (", mags_string, " MAGs respectively).")` Host families most often targeted within `r df_filtered[1,1]` include `r fam_string_join(host_fam_count,df_filtered[1,1])[[1]]` (`r fam_string_join(host_fam_count,df_filtered[1,1])[[2]]` MAGs respectively). The host family most often targeted within `r df_filtered[2,1]` is `r fam_string_join(host_fam_count,df_filtered[2,1])[[1]]` (`r fam_string_join(host_fam_count,df_filtered[2,1])[[2]]` MAGs respectively). The host family most often targeted within `r df_filtered[3,1]` is `r fam_string_join(host_fam_count,df_filtered[3,1])[[1]]` (`r fam_string_join(host_fam_count,df_filtered[3,1])[[2]]` MAGs). The host family most often targeted within `r df_filtered[4,1]` is  `r fam_string_join(host_fam_count,df_filtered[4,1])[[1]]` (`r fam_string_join(host_fam_count,df_filtered[4,1])[[2]]` MAGs). The host family most often targeted within `r df_filtered[5,1]` is  `r fam_string_join(host_fam_count,df_filtered[5,1])[[1]]` (`r fam_string_join(host_fam_count,df_filtered[5,1])[[2]]` MAGs).

```{r,include=FALSE}
# SingleM reps in p-h pairing

#MAG representatives of the singleM indicators
v_mag_indicator_reps<-fread("../../data/results/tables/mag_motu_indicator_reps_pha.csv")%>%
  rename("OTU_ID"="sequence")%>%
  select(OTU_ID,MAG,clean_MAG:species,`Indicator taxa`,vOTU,viral_tax,lifestyle,iphop,phase,host)%>%
  unique%>%
  mutate(ph_prediction=case_when(iphop=="TRUE"&phase=="TRUE"~"both",
                                 iphop=="TRUE"&phase=="FALSE"~"iphop",
                                 iphop=="FALSE"&phase=="TRUE"~"phase"))%>%
  select(-iphop,-phase)%>%
  as.data.frame

v_mag_indicator_reps_agger<- unique(v_mag_indicator_reps[,c("clean_MAG","OTU_ID","Indicator taxa","host")])%>%
  group_by(clean_MAG)%>%
  summarise(n_mOTU_ID=n_distinct(OTU_ID),
            n_indic_taxa=n_distinct(`Indicator taxa`),
            indic_taxa=paste(unique(`Indicator taxa`),collapse = " - "),
            n_bat_host=n_distinct(host),
            bat_host=paste(unique(host),collapse = " - "))

indicator_species_mag_count<-unique(v_mag_indicator_reps[,c("clean_MAG","OTU_ID","Indicator taxa","host")])%>%
  aggregate(cbind(`Indicator taxa`,host)~clean_MAG, function(x)paste(unique(x), collapse = " - "))
unique(v_mag_indicator_reps[v_mag_indicator_reps$iphop==""])

# Characterizing the taxa of the indicator MAGs
in_mag_taxa <- unique(v_mag_indicator_reps[, c("MAG", "phylum","host_family","genus")]) %>% 
  mutate(num_mags=1)%>%
  aggregate(num_mags ~ phylum+host_family+genus, sum) %>%
  arrange(desc(num_mags))

#Function for genus counting
gen_string_join<-function(df,phy){
  phy_select<-df%>%
    filter(phylum==phy)
  fam_join<-join_with_and(phy_select$genus)
  num_join<-join_with_and(phy_select$num_mags)
  return(list(fam_join,num_join))
}
 #Phylum count
in_mag_phy_count <- unique(v_mag_indicator_reps[, c("MAG", "phylum")]) %>% 
  mutate(num_mags=1)%>%
  aggregate(num_mags ~ phylum, sum) %>%
  arrange(desc(num_mags))

gen_string_join(in_mag_taxa,in_mag_phy_count[1,1])
```

  Based on the singleM genus-level recovery estimates, we were able to determine that `r round(length(unique(v_mag_indicator_reps$OTU_ID))/sum(ind_sp$total_mOTUs)*100)`% (n=`r length(unique(v_mag_indicator_reps$OTU_ID))`) of the `r sum(ind_sp$total_mOTUs)` mOTU indicator species are represented by `r round(length(unique(v_mag_indicator_reps$MAG))/length(unique(ip_ph_compare$MAG))*100)`% (n=`r length(unique(v_mag_indicator_reps$MAG))`) of the `r length(unique(ip_ph_compare$MAG))` MAGs that were paired to phages (Fig. 6). `r in_mag_phy_count[1,2]` of these MAGs belong to the phylum `r in_mag_phy_count[1,1]` (in the genre `r gen_string_join(in_mag_taxa,in_mag_phy_count[1,1])[[1]]` (`r gen_string_join(in_mag_taxa,in_mag_phy_count[1,1])[[2]]` respectively)), `r in_mag_phy_count[2,2]` belong to the phylum `r in_mag_phy_count[2,1]` (in the genre `r gen_string_join(in_mag_taxa,in_mag_phy_count[2,1])[[1]]` (`r gen_string_join(in_mag_taxa,in_mag_phy_count[2,1])[[2]]` respectively)), and `r in_mag_phy_count[3,2]` belong to the phylum `r in_mag_phy_count[3,1]` (in the genre `r gen_string_join(in_mag_taxa,in_mag_phy_count[3,1])[[1]]` (`r gen_string_join(in_mag_taxa,in_mag_phy_count[3,1])[[2]]` respectively)).

   `r nrow(indicator_species_mag_count[indicator_species_mag_count$host=="Mimon crenulatum",])` MAGs were indicators for _P. hastatus_, `r nrow(indicator_species_mag_count[indicator_species_mag_count$host=="Phyllostomus hastatus",])` were indicators for _M. crenulatum_, and `r nrow(indicator_species_mag_count[indicator_species_mag_count$host=="Mimon crenulatum - Phyllostomus hastatus",])` was an indicator for both _P. hastatus_ and _M. crenulatum_.

  These `r length(unique(v_mag_indicator_reps$MAG))` indicator MAGs were paired with `r round(length(unique(v_mag_indicator_reps$vOTU))/length(unique(ip_ph_compare$vOTU))*100)`% (n=`r length(unique(v_mag_indicator_reps$vOTU))`) of the `r length(unique(ip_ph_compare$vOTU))` phage paired to host MAGs. `r nrow(unique(v_mag_indicator_reps[, c("MAG", "ph_prediction")])[v_mag_indicator_reps[!duplicated(v_mag_indicator_reps[, c("MAG", "ph_prediction")]), "ph_prediction"] == "phase", ])` phase-host prediction involving indicator MAGs were made with Hi-C, `r nrow(unique(v_mag_indicator_reps[, c("MAG", "ph_prediction")])[v_mag_indicator_reps[!duplicated(v_mag_indicator_reps[, c("MAG", "ph_prediction")]), "ph_prediction"] == "iphop", ])` were made by IPHoP, and `r nrow(unique(v_mag_indicator_reps[, c("MAG", "ph_prediction")])[v_mag_indicator_reps[!duplicated(v_mag_indicator_reps[, c("MAG", "ph_prediction")]), "ph_prediction"] == "both", ])` were made by both the Hi-C and IPHoP methods. `r nrow(unique(v_mag_indicator_reps[,c("vOTU","lifestyle")])[unique(v_mag_indicator_reps[,c("vOTU","lifestyle")])$lifestyle=="lytic",])` of vOTUs matched to indicator MAGs were predicted to be lytic, `r nrow(unique(v_mag_indicator_reps[,c("vOTU","lifestyle")])[unique(v_mag_indicator_reps[,c("vOTU","lifestyle")])$lifestyle=="lysogenic",])` were lysogenic, while the lifestyle of `r nrow(unique(v_mag_indicator_reps[,c("vOTU","lifestyle")])[unique(v_mag_indicator_reps[,c("vOTU","lifestyle")])$lifestyle=="unclear",])` was unable to be predicted.

```{r Fig6, echo = FALSE, message=FALSE, fig.cap='**Figure 6. Phage-host network analysis.**'}

knitr::include_graphics("../../data/results/figures/phage-host_bat_network_final_04232025.pdf")
```

## AMR Genes
  We identified a total of **268** unique AMR gene variants within the MAGs and vMAGs across the three bat species. **27** AMR genes were shared between the MAGs and vMAGs.

### MAG resistome diversity
  Across the gut MAG associated resistome of the 3 bat species, Macaroline resistance were the most abundant AMR class in both _M. crenulatum_ and _P. discolor_ while Tetracycline were the most abundant in _P. hastatus_. The second most prevalent AMR varied by bat species, with Tertracycline resistance observed in _M. crenulatum_, Glycopeptide resistance observed in _P. discolor_, and Macrolide resistance abserved in _P. hastatus_ (**Fig X**). The most abundant gene associated with Macrolide resistance was abc-f (all three bat species), _mef(B)_ ( _P. hastatus_ ), _erm_ ( _M. crenulatum_ ), and _estT_ ( _M. crenulatum_ and _P. discolor_). Tetracycline resistance was dominated by _tet_ genes ( _tet_, _tetA(46)_, and _tetB(46)_) within all three bat species while _mepA_ was more abundant in both _P. hastatus_ and _P. discolor_. The most predominate genes associated with Glycopeptide resistance was represented _vanR_, _vanS_, _vanH_, and _vanR-B_ in the guts across all three bat species while _vanY-D_ was highly abundant in the guts of _M. crenulatum_ and _P. discolor_.

  Within the MAGs, a total of **52** AMR genes across **X** AMR classes were unique to _M. crenulatum_, **22** AMR genes across **X** AMR classes were unique to _P. hastatus_, while no AMR genes were found to be unique to _P. discolor_ (**Fig X**). Unique AMR genes within _M. crenulatum_ were within the classes **X,Y,andZ**. The most abundant of these were **A,B,andC**. The **22** AMR genes unique to  _P. hastatus_ were within the classes **X,Y,andZ** with the most predominate genes being **A,B,andC**. Interesting **71** AMR genes (representing **X** classes) were shared between _M. crenulatum_ and _P. hastatus_ while only 3 (representing **X** classes) were shared between _P. discolor_ and _P. hastatus_ and only 2 between (representing **X** classes) _M. crenulatum_ and _P. discolor_.

### vMAG resistome diversity

# Discussion
  Our novel study is the first to offer empirical evidence of phage-host interactions in the gut microbiome of bats using metagenomics in conjunction with Hi-C techniques. (**Talk about previous attempts of phage-host studies in bat guts and the methods they used**).

  (**Need an introduction to this paragraph here that leads to the hypothesis in the following sentence**). Therefore, we hypothesized that the bat gut virome acts as a phage-mediated immune system against pathogenic prokaryotes. In support of this hypothesis, those prokaryotic families that had more frequent pairings with phage were those that contained known pathogenic members. For example, the families `r fam_string_join(host_fam_count,df_filtered[1,1])[[1]]` within the bacteria phylum `r df_filtered[1,1]`, the family `r fam_string_join(host_fam_count,df_filtered[2,1])[[1]]` within the bacteria phylum `r df_filtered[2,1]`, the and family `r fam_string_join(host_fam_count,df_filtered[3,1])[[1]]` within the bacteria phylum `r df_filtered[3,1]` all contain members of known human pathogens (**Citations**).

  While the family `r fam_string_join(host_fam_count,df_filtered[4,1])[[1]]` within the bacteria phylum `r df_filtered[4,1]` contain no known pathogenic members, members are associated with the human gut (**Citation**).

Macro Eukarotic 
* (AND) _Iridoviridae_
* (AND) _Bornaviridae_ 
* (AND) _Paroviruses_
Micro Eukarotic 
* (AND) _Mimiviridae_
Bacterial phage 
* (AND) an unclassified family of _Kyanoviridae_ 
* (AND) _Inoviridae_ 
* (AND) an unclassified family of _Herelleviridae_

# Conclusion

# Acknowledgement

### 2025-06-30 phasegenome_out/doc/paper/manuscript.tex
Manuscript update

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{helvet}
\renewcommand*\familydefault{\sfdefault}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\section{Resolving gut microbiome networks within
Chiropterans}\label{resolving-gut-microbiome-networks-within-chiropterans}

Timothy J. Rogers, Laurel R. Yohe, Richard A. White \({\dagger}\)

Department of Bioinformatics, University of North Carolina, Charlotte,
NC 28223; Department of Bioinformatics and Genomics, North Carolina
Research Center (NCRC), Kannapolis, NC

\({\dagger}\) To whom correspondence should be addressed

\newpage

\subsection{Abstract (250 words)}\label{abstract-250-words}

\newpage

\subsection{Introduction}\label{introduction}

\begin{itemize}
\tightlist
\item
  Bats are known carriers of human associated pathogens
\item
  The reason bats are bioreactors is not understood
\item
  The diet of bats may contribute to the gut microbiota makeup
\item
  Phage associated with these microbiota can benifit and hinder
  microbial populations and have an impack on bat immune responses
\item
  Some ideas suggest viral tolerance is linked to

  \begin{itemize}
  \tightlist
  \item
    Uniqueness of bats and their variation in Diets

    \begin{itemize}
    \tightlist
    \item
      Diversity
    \item
      Ecological role
    \item
      Pathogenic role
    \end{itemize}
  \end{itemize}
\item
  Methods for characterizing microbiome and virome. As well as methods
  for linking the two

  \begin{itemize}
  \tightlist
  \item
    Culture dependant vs independent

    \begin{itemize}
    \tightlist
    \item
      What has been discovered with these methods
    \item
      How these methods have been applied to bats

      \begin{itemize}
      \tightlist
      \item
        What has been found
      \item
        What has yet to be discribed
      \end{itemize}
    \end{itemize}
  \end{itemize}
\item
  Goal of this study
\end{itemize}

\subsection{Materials and Methods}\label{materials-and-methods}

\begin{itemize}
\tightlist
\item
  Sample location discription
\item
  Sample collection
\item
  Sample processing \#\#\# Phase genomics portion
\item
  (AND)
\end{itemize}

\subsubsection{In house analysis of MAGs, vcontigs, and viral-host
pairing}\label{in-house-analysis-of-mags-vcontigs-and-viral-host-pairing}

\paragraph{vOTU curration}\label{votu-curration}

\begin{itemize}
\tightlist
\item
  (AND) Genomad (v x.x.x) was used to varify contigs identified as viral
  by phase genomics and for taxonomic identification.
\item
  (AND) We also ran all assembly files from phase genome through genomad
  to identify potential viral sequences that the phase genome pipeline
  may have missed
\item
  (AND) Quality filtered viral sequences were then clustered into
  species-level equivalent viral operational taxanomic units (vOTUs) at
  95\% average nucleatide identity over 85\% of the alignment fraction
  of the shorter sequence using the greedy centroid algorithm
  (anicalc.py and aniclust.py) from checkV (v x.x.x).
\end{itemize}

\paragraph{MAG curration}\label{mag-curration}

\begin{itemize}
\tightlist
\item
  (AND) MAGs from phase metagenomic were quality filtered using checkM
  (v x.x.x).
\item
  (AND) Quality filtered MAGs were then dereplicated using dRep (v
  x.x.x) at 99\% ANI with the following settings (--S\_algorithm
  fastANI, -comp 50, --SkipMash)
\item
  (AND) MAGs representatives were taxonomically identified using the
  Genome Taxonomy Database Toolkit (GTDB-Tk v x.x.x)
\end{itemize}

\paragraph{vOTU and MAG coverage}\label{votu-and-mag-coverage}

\begin{itemize}
\tightlist
\item
  (AND) Quality controled reads were mapped to the vOTUs and
  representative MAGs using bowtie (v x.x.x) and sam files were
  converted to indexed and sorted bam files using samtools (v x.x.x).
\item
  (AND) Bam files were then fed into anvi'o (v x.x.x) to calculate the
  Q2Q3 coverage of vOTUs and MAGs across bat species.
\item
  (AND) Q2Q3 coverages for both vOTUs and MAGs were then normalized by
  sequence length.
\item
  (AND) For alpha diveristy analyzes, normalized Q2Q3 coverages for both
  vOTUs and MAGs were rarefied using the rrarify function of the vengan
  R package.
\item
  (AND) For beta diversity analyzes, normalized Q2Q3 coverages for both
  vOTUs and MAGs were used to create rarefied Hellinger distance
  matrices using the vegan (v x.x.x) and labdsv (v x.x.x) R packages.
\item
  (AND) For abundance pattern analyzes, normalized Q2Q3 coverages were
  coverted into units of GCPM (Genome Copy Per Million Reads) for MAGs,
  as laid out in Rogers et al.~2022 (citation), and TPM (Transcripts Per
  Million Reads) for vOTUs.
\item
  (AND) Abundance correlations of vOTUs and MAGs across and within
  samples were determined using hierarchical clustering of Spearman rank
  correlation distance matrice in base R (v x.x.x).
\end{itemize}

\paragraph{Metabolic predictions}\label{metabolic-predictions}

\begin{itemize}
\tightlist
\item
  The program metacerberus (v x.x.x) was used for gene anotation of both
  viral sequences and MAGs using the following data bases: Functional
  Ontology Assignments for Metagenomes (FOAM), KEGG, CAZy/dbCAN, VOG,
  pVOG, PHROG, and COG.
\end{itemize}

\paragraph{AMR and CAZY gene abundance
patterns}\label{amr-and-cazy-gene-abundance-patterns}

\begin{itemize}
\tightlist
\item
  (AND) Read coverage and abundance for AMR and CAZY genes were
  calculated using a custom ihhouse script.
\end{itemize}

\paragraph{Linking virus to host}\label{linking-virus-to-host}

\begin{itemize}
\tightlist
\item
  (AND) IPHoP (v x.x.x) was used to link all viral sequences to all
  medium to high quality MAGs, not just to the representative vOTUs and
  MAGs.
\item
  (AND) The viral and MAG IDs of those within the viral-host predictions
  from both IPHoP and Phase genomes were replaced with the corresponding
  representative vOTU and MAG IDs.
\end{itemize}

\paragraph{Statistical analysis}\label{statistical-analysis}

\begin{itemize}
\tightlist
\item
  (AND) Differences in MAG and vOTU composition among bat species were
  analyzed via distance based redundancy analysis (db-RDA) on a
  quntitative Hellinger distance matirx.
\item
  (AND) The richness and evenness of MAGs and vOTUs were compared using
  the R package microbiome for richness (chao1) and the package hillR (v
  x.x.x) for evenness (Pielou's index).
\item
  (AND) Statistical significants was determined based on 9999
  permutations of the data in the vegan R package.
\end{itemize}

\subsection{Results}\label{results}

\subsubsection{MAGs}\label{mags}

\begin{itemize}
\item
  (AND) We recovered 5 Archaeal and 182 Bacterial medium to high quality
  MAGs with \(\geq\) 50 \% and \(\leq\) 9 \% redundancy.
\item
  (AND) Of these 187 MAGs, 106 were \(\geq\) 90\% complete and \(\leq\)
  10\% redundant, while YY are of high quality as determined by
  parameters layed out in XXX et al.~(** TABLE **).
\item
  (AND) \emph{Bacillota}, \emph{Actinomycetota}, and
  \emph{Pseudomonadata} were the predominate phylum across the gut
  microbiome within all 3 of the bat species (\textbf{Figure}).
\item
  (AND) No significant differnces were found in the richness (Choa1) or
  evenness (Pielou's Index) of prokaryotic MAGs across the 3 bat
  species.
\item
  (AND) db-RDA analysis revealed no significant differnce in the
  \(B\)-diversity of the prokaryotic MAGs across the 3 bat species.
\end{itemize}

\begin{figure}
\includegraphics[width=400px,height=350px]{../../data/results/figures/FigX_top10_most_abun_mag_taxa} \caption{Distribution of the top 10 most abundant prokaryotic families across the gut microbiome of the three bat species}\label{fig:FigX}
\end{figure}

\subsubsection{vOTUs}\label{votus}

\begin{itemize}
\tightlist
\item
  (AND) A total of 17702 potential viral sequences were ID from the 6
  individual assemblies.
\item
  (AND) Clustering at 95\% ANI over 85\% of the shortest sequence
  identified 16289 viral operational taxanomic units (vOTUs) that were
  \(\geq\) 1 kbp in length.
\item
  (AND) Further filtering for sequences \(\geq\) 2.5 kpb resulted in a
  final set of 6235 vOTUs.
\item
  (AND) After rarefaction, 5582 vOTUs were retained.
\item
  (AND) CheckV was used to assess the quality of these sequences,
  revealing that 361 (6\%) were \(\geq\) 50\% complete including 39
  complete vOTUs that identified on the bases of direct terminal repeats
  (DTR), 74 high quality vOTUs that were identified on the bases of AAI
  (54 vOTUs) and HMM (20 vOTUs), 198 medium quality vOTUs that were
  idenified on the bases on AAI (136 vOTUs) and HMM (62 vOTUs), and 50
  low quality vOTUs that were identified based on AAI (41 vOTUs) and HMM
  (9 vOTUs). The reset of the vOTUs (5221) were of low quality (3298) or
  the quality was undetermined (1923).
\item
  (AND) An unclassified order of the class \emph{Caudoviricetes},
  families \emph{Retroviridae}, \emph{Adintoviridae}, and
  \emph{Iridoviridae}, an unclassified family of \emph{Kyanoviridae},
  families \emph{Inoviridae} and \emph{Bornaviridae}, an unclassified
  family of \emph{Herelleviridae}, families \emph{Mimiviridae} and
  \emph{Parvoviridae} were the most predominate viral taxa across the
  gut virome with all 3 of the bat species (\textbf{Figure}).
\item
  (But) No statistically significant differences were found in the viral
  richness, envenness, or \(B\)-diversity nor did indicator analysis
  reveal any indicator viral spices for the bat species.
\end{itemize}

\begin{figure}
\includegraphics[width=400px,height=350px]{../../data/results/figures/FigX_top10_most_abun_viral_taxa} \caption{Distribution of the top 10 most abundant viral families across the gut virome of the three bat species}\label{fig:FigY}
\end{figure}

\paragraph{Metabolic capabilities}\label{metabolic-capabilities}

\begin{itemize}
\tightlist
\item
  (AND)
\end{itemize}

\subsubsection{Virus-host predicitons}\label{virus-host-predicitons}

\begin{itemize}
\tightlist
\item
  (AND) A total of 1308 viru-host prodictions were made using phase HiC
  and IPHoP methods.
\item
  (AND) These predictions included 186 MAGs and 953 vOTUs.
\item
  (AND) Of these, 352 were made by only phase, 867 were made by only
  IPHoP, and 80 were agreed matches between both methods
  (\textbf{Figure}).
\item
  (AND) Of those predictions made by IPHoP, 565 are based on blastn and
  391 are based on iPHoP-RF
\item
  (AND) MAGs within the bacteria phyla \emph{Actinomycetota},
  \emph{Bacillota}, \emph{Pseudomonadota}, and \emph{Desulfobacterota}
  had the highest frequency of being targeted by viruses (84, 62, 18,
  and 13 MAGs respectively).
\item
  (AND) Host families most often targeted within \emph{Actinomycetota}
  include \emph{Mycobacteriaceae} (20 MAGs) and \emph{Nocardiaceae} (10
  MAGs).
\item
  (AND) The host family most often targeted within \emph{Bacillota} is
  \emph{Enterococcaceae} (15 MAGs).
\item
  (AND) The host family most often targeted within \emph{Pseudomonadota}
  is \emph{Enterobacteriaceae} (6 MAGs).
\item
  (AND) The host family most often targeted within
  \emph{Desulfovibrionaceae} is \emph{Desulfobacterota} (9 MAGs).
\end{itemize}

\begin{figure}
\includegraphics[width=400px,height=350px]{../../data/results/figures/updated_venndiagram_phase_iphop} \caption{Venn diagram showing the number of virus-host predictions made by each method used}\label{fig:FigZ}
\end{figure}

\begin{figure}
\includegraphics[width=400px,height=350px]{../../data/results/figures/Bat_gut_viral_host_network_figure} \caption{Virus-host network analysis}\label{fig:FigA}
\end{figure}

\subsection{Descusion}\label{descusion}

\begin{itemize}
\tightlist
\item
  (AND) The families \emph{Mycobacteriaceae} and \emph{Nocardiaceae}
  within the bacteria phylum \emph{Actinomycetota} contain known
  pathogenic members (\textbf{CITATION}).
\item
  (AND) The family \emph{Enterococcaceae} within the bacteria phylum
  \emph{Bacillota} contains known pathogenic members
  (\textbf{CITATION}).
\item
  (AND) The family \emph{Enterobacteriaceae} within the bacteria phylum
  \emph{Pseudomonadota} contain known pathogenic memebers
  (\textbf{CITATION}).
\item
  (AND) The family \emph{Desulfovibrionaceae} within the bacteria phylum
  \emph{Desulfobacterota} contain known opportunistic pathogenic members
  (\textbf{CITATION}).
\end{itemize}

Macro Eukarotic * \emph{Iridoviridae} \emph{\emph{Bornaviridae}
}\_Paroviruses\_

Micro Eukarotic *\emph{Mimiviridae}

Bacterial phage * an unclassified family of \emph{Kyanoviridae} *
\emph{Inoviridae} * an unclassified family of \emph{Herelleviridae} \#\#
Conclusion

\subsection{Acknowledgement}\label{acknowledgement}

\end{document}

### 2025-06-30 phasegenome_out/doc/paper/test.Rmd
Manuscript update

--- 2025-03-03 phasegenome_out/doc/paper/manuscript.rmd
@@
 ---
-title: Resolving gut microbiome networks within Chiropterans
-authors:
-  - name: Timothy J. Rogers
-    department: Department of Bioinformatics
-    affiliation: University of North Carolina
-    location: Charlotte, NC 28223
-    email: troger50@charlotte.edu
-  - name: Laurel R. Yohe
-    department: Department of Bioinformatics and Genomics
-    affiliation: North Carolina Research Center (NCRC)
-    location:  Kannapolis, NC 28081
-    email:  lyohe1@charlotte.edu
-  - name: Richard A. White ${^\dagger}$
-    department: Department of Bioinformatics
-    affiliation: University of North Carolina
-    location: Charlotte, NC 28223
-    email: rwhit101@charlotte.edu
-abstract: |
-  Enter the text of your abstract here.
-keywords:
-  - Chiropterans
-  - microbiome
-  - virome
-  - metagenomics
-bibliography: "`r rbbt::bbt_write_bib('references.bib', overwrite = TRUE)`" 
-#bibliography: references.bib
-biblio-style: unsrt
-output: rticles::arxiv_article
-geometry: margin=1.0in
-font-size: 11pt
-header-include:
-  -\usepapackage(helvet)
-  -\renewcommand*\familydefault{\sfdefault}
+title: "test"
+output: html_document
+date: "2025-03-03"
+---
 
----
 ${\dagger}$ To whom correspondence should be addressed
@@
 ### MAGs
-```{r, echo=FLASE}
+```{r, echo=FALSE}
 # Quality and length of rarefied MAGs
-mag_stats<-fread("/projects/raw_lab/projects/Bats/bat_gut_micro/phasegenome_out/data/results/tables/mag_stats.csv")
+mag_stats<-fread("../../data/results/tables/mag_stats.csv")
 # Abudance information:
-mag_fam_rel_abun<-fread("/projects/raw_lab/projects/Bats/bat_gut_micro/phasegenome_out/data/results/tables/mag_fam_rel_abun.csv")
+mag_fam_rel_abun<-fread("../../data/results/tables/mag_fam_rel_abun.csv")
 phylum_rel_abun<-mag_fam_rel_abun%>%select(-V1,-class:-family)%>%
   aggregate(.~phylum,sum)
-colSums(phylum_rel_abun[,-1])
+#colSums(phylum_rel_abun[,-1])
 ```
@@
 * (AND) db-RDA analysis revealed no significant differnce in the $B$-diversity of the prokaryotic MAGs across the 3 bat species.
-\includegraphics{../data/results/figures/FigX_top10_most_abun_viral_taxa.pdf}
+
 ### vOTUs

### 2025-06-30 phasegenome_out/doc/paper/draft_manuscript.rmd
Manuscript update
Manuscript Update

@@
 ### MAG resistome diversity
-  Across the gut MAG associated resistome of the 3 bat species, Macaroline resistance were the most abundant AMR class in both _M.
+  Across the gut MAG associated resistome of the 3 bat species, Macaroline resistance was the most abundant AMR class in both _M.
 crenulatum_ and _P.
@@
 
+```{r Fig7, echo = FALSE, message=FALSE, fig.cap='**Figure 7. Venn diagram of MAG associated AMR genes.**'}
+
+knitr::include_graphics("../../data/results/AMR_Venn_diet.pdf")
+```
+
 ### vMAG resistome diversity
+  Within the gut vMAG associated resistome of the 3 bat species, Macaroline resistance was the most abundant AMR class in both _M.
+crenulatum_ and _P.
+discolor_.
+In fact, this was the only class of antimicrobial resistance found within _P.
+discolor_.
+The Mupirocin resistance class was the most abundant in _P.
+hastatus_.
+The second most prevalent AMR varied between _M.
+crenulatum_ and _P.
+hastatus_, with Quinolone resistance observed in _M.
+crenulatum_ and  Macaroline resistance observed in _P.
+hastatus_ (**Fig X**).
+The only AMR gene associated with Macrolide resistance was _abc-f_ (all three bat species).
+Quinolone resistance was represented by _qnrVC_, _qnrE_, _qnrS_, _qnrA_, and _qnrB_ within _M.
+crenulatum_ and by qnrD in _P.
+hastatus_.
 
+  A total of **11** AMR genes across **X** AMR classes were unique to _M.
+crenulatum_, **5** AMR genes across **X** AMR classes were unique to _P.
+hastatus_, while no AMR genes were found to be unique to _P.
+discolor_ (**Fig X**).
+Unique AMR genes within _M.
+crenulatum_ were within the classes **X,Y,andZ**.
+The most abundant of these were **A,B,andC**.
+The **5** AMR genes unique to  _P.
+hastatus_ were within the classes **X,Y,andZ** with the most predominate genes being **A,B,andC**.
+Interesting **7** AMR genes (representing **X** classes) were shared between _M.
+crenulatum_ and _P.
+hastatus_ while none were shared between _P.
+discolor_ and _P.
+hastatus_ nor between _M.
+crenulatum_ and _P.
+discolor_.
+
+```{r Fig8, echo = FALSE, message=FALSE, fig.cap='**Figure 8. Venn diagram of vMAG associated AMR genes.**'}
+
+knitr::include_graphics("../../data/results/figures/vcon_amr_venn.pdf")
+```
 # Discussion

### 2025-07-07 phasegenome_out/doc/paper/draft_manuscript.rmd
update

@@
   Results were imported into R for analysis and visualization using the R packages data.table (**CITATION**), tidyverse (**CITATION**), ggplot (**CITATION**), vegan (**CITATION**), hillR (**CITATION**), phyloseq (**CITATION**), ggtext (**CITATION**), and ggpubr (**CITATION**).
-mOTUs identified by SingleM were used to assess alpha and beta diversity.
+mOTUs identified by SingleM were used to assess $a$- and β-diversity.
 Differences in prokaryotic community composition among bat species were analyzed using distance-based redundancy analysis (db-RDA) on a Hellinger distance matrix.
@@
 #Calling in phyloseq object
-rel_abun_phy<-readRDS("../../data/results/tables/singleM__rarefied_phy.rds")
+rel_abun_phy<-readRDS("../../data/results/singleM__rarefied_phy.rds")
 
@@
 An unclassified order of the class _Caudoviricetes_, families _Retroviridae_, _Adintoviridae_, and _Iridoviridae_, an unclassified family of _Kyanoviridae_, families _Inoviridae_ and _Bornaviridae_, an unclassified family of _Herelleviridae_, families _Mimiviridae_ and _Parvoviridae_ were the most predominate viral taxa across the gut virome with all 3 of the bat species (**Figure 4**).
-No statistically significant differences were found in the viral richness, envenness, $\alpha$-diversity, or $B$-diversity nor did indicator analysis reveal any indicator viral spices for the bat species.
+No statistically significant differences were found in the viral richness, envenness, $\alpha$-diversity, or β-diversity nor did indicator analysis reveal any indicator viral spices for the bat species.
 
@@
 discolor_, and Macrolide resistance abserved in _P.
-hastatus_ (**Fig X**).
+hastatus_ (**Fig 7**).
 The most abundant gene associated with Macrolide resistance was abc-f (all three bat species), _mef(B)_ ( _P.
@@
 
-  Within the MAGs, a total of **52** AMR genes across **X** AMR classes were unique to _M.
-crenulatum_, **22** AMR genes across **X** AMR classes were unique to _P.
+  Within the MAGs, a total of **52** AMR genes across **15** AMR classes were unique to _M.
+crenulatum_, and **22** AMR genes across **8** AMR classes were unique to _P.
 hastatus_, while no AMR genes were found to be unique to _P.
-discolor_ (**Fig X**).
-Unique AMR genes within _M.
-crenulatum_ were within the classes **X,Y,andZ**.
-The most abundant of these were **A,B,andC**.
-The **22** AMR genes unique to  _P.
-hastatus_ were within the classes **X,Y,andZ** with the most predominate genes being **A,B,andC**.
-Interesting **71** AMR genes (representing **X** classes) were shared between _M.
-crenulatum_ and _P.
-hastatus_ while only 3 (representing **X** classes) were shared between _P.
-discolor_ and _P.
-hastatus_ and only 2 between (representing **X** classes) _M.
-crenulatum_ and _P.
-discolor_.
+discolor_ (**Fig.
+X**).
+Although these AMR genes were unique to either _M.
+crenulatum_ or _P.
+hastatus_, **7** of the **8** AMR gene classes present in _P.
+hastatus_ (Aminoglycoside, β-lactam, Trimethoprim, Macrolide, Fosfomycin, Lincosamide, and Colistin resistance) were also found in _M.
+crenulatum_.
+Only the Streptothricin class was unique to _P.
+hastatus_.
+AMR classes unique to _M.
+crenulatum_ included Bleomycin, Phenicol, Lincosamide/Macrolide, Phenicol/Oxazolidinone, Sulfonamide, Tetracycline, Glycopeptide, and Streptogramin resistance.
+
+  Among the unique AMR genes, the most prominent in _M.
+crenulatum_ were _acc(2')-lc_, _aac(6')-Il_ (Aminoglycoside), _dfrF_ (Trimethoprim), _blaDHA_, _blaGMB_, _blaMAB_, _blaBUT_, _blaOCH_, _cfiA_ (β-lactam), and _vanZ-Pt_ (Glycopeptide).
+In contrast, the most predominant AMR genes in _P.
+hastatus_ were _aac(3)-VIII_, _aac(6')-I_, _aac(6')-Ian_, _aacA10_, _aph(3')-V_, _aph(7')-Ia_, _aph(9)-Ia_ (Aminoglycoside), _dfrA36_, _dfrG_ (Trimethoprim), _blaPDC_, _blaZ_ (β-lactam), _mcr-10_ (Colistin), _fosC_ (Fosfomycin), _erm(55)_, _mph(C)_ (Macrolide), and _sat2_ (Streptothricin).
+All other AMR genes had abundances of **< 5** TPM.
 
@@
 crenulatum_ and  Macaroline resistance observed in _P.
-hastatus_ (**Fig X**).
+hastatus_ (**Fig 8**).
 The only AMR gene associated with Macrolide resistance was _abc-f_ (all three bat species).
@@
 
-  A total of **11** AMR genes across **X** AMR classes were unique to _M.
-crenulatum_, **5** AMR genes across **X** AMR classes were unique to _P.
+  A total of **11** AMR genes across **6** AMR classes were unique to _M.
+crenulatum_, **5** AMR genes across **5** AMR classes were unique to _P.
 hastatus_, while no AMR genes were found to be unique to _P.
 discolor_ (**Fig X**).
-Unique AMR genes within _M.
-crenulatum_ were within the classes **X,Y,andZ**.
-The most abundant of these were **A,B,andC**.
-The **5** AMR genes unique to  _P.
-hastatus_ were within the classes **X,Y,andZ** with the most predominate genes being **A,B,andC**.
-Interesting **7** AMR genes (representing **X** classes) were shared between _M.
-crenulatum_ and _P.
-hastatus_ while none were shared between _P.
-discolor_ and _P.
-hastatus_ nor between _M.
-crenulatum_ and _P.
-discolor_.
+While these AMR genes are unique to either _M.
+crenulatum_ or _P.
+hastatus_, Quinolone, Trimethoprim, and Glycopeptide resistant classes are present in both bat species.
+AMR classes Phenicol, Rifamycin, and Streptogramin are unique classes within _M.
+crenulatum_ while Mupirocin and Thiostrepton are unique AMR classes within _P.
+hastatus_.
+
+  Among the unique AMR genes, the most prominent in _M.
+crenulatum_ were _qnrA_, _qnrB_, _qnrVC_, _qnrE_, _qnrS_ (Quinolone), _dfrA_ (Trimethoprim), and _vat_ (Streptogramin).
+In contrast, the most predominant AMR genes in _P.
+hastatus_ were _mupA_ (Mupirocin), _dfra35_ (Trimthoprim), _nshR_ (Thiostrepton), _vanY_ (Glysopeptide), and _qnrD_ (Quinolone).
+All other AMR genes had abundances of **< 5** TPM.
 

### 2025-07-07 phasegenome_out/doc/paper/draft_manuscript.rmd
update

@@
 ```
+
+F
 
@@
 
-```{r Fig6, echo = FALSE, message=FALSE, fig.cap='**Figure 6. Phage-host network analysis.**'}
-
-knitr::include_graphics("../../data/results/figures/phage-host_bat_network_final_04232025.pdf")
-```
+# ```{r Fig6, echo = FALSE, message=FALSE, fig.cap='**Figure 6. Phage-host network analysis.**'}
+# 
+# knitr::include_graphics("../../data/results/figures/phage-host_bat_network_final_04232025.pdf")
+# ```
 
